### 2015-11-16 test_2.tex

\documentclass[crop,tikz]{standalone}

\usepackage{pgfplots}
\usepackage[separate-uncertainty=true,detect-mode=true]{siunitx}

\begin{document}

    \begin{tikzpicture}
        \begin{axis}[
            view={0}{90},
            xlabel=quantity of wine,
            ylabel=grape quality,
            colorbar,
            colormap/blackwhite
            ]

            \addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0,0,   1) (1,0,   1) (1,1.25,1) (0,1.25, 1)};
            \addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0,1.25,1) (1,1.25,0) (1,2,   0) (0,2,   0.2)};
            \addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0,2,   1) (1,2,   0) (1,3,   0) (0,3,   0.3)};

            \addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(1,0,   0) (2.5,0,   0) (2.5,1.25,0) (1,1.25,0)};
            \addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(1,1.25,1) (2.5,1.25,0) (2.5,2,   0) (1,2,   0.25)};
            \addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(1,2,   1) (2.5,2,   0) (2.5,3,   0) (1,3,   0.05)};

            \addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(2.5,0,   1) (3,0,   0) (3,1.25,0) (2.5,1.25,0.25)};
            \addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(2.5,1.25,1) (3,1.25,0) (3,2,   0) (2.5,2,   0.1)};
            \addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(2.5,2,   1) (3,2,   0) (3,3,   0) (2.5,3,   0.3)};
        \end{axis}
    \end{tikzpicture}

    % \label{fig:plot_label}
    % \caption{Probability density under MODL discretization for two variables from the Wine dataset}
\end{document}