### 2015-11-17 graph_housing_exp1_distr_3_5.tex
Experiment part

	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			enlargelimits=0,
			zmin=0.0, zmax=1.0,
			xlabel=C,
			ylabel=E,
			colorbar,
			point meta min=0, point meta max=0.5,
			colormap={}{ gray(0cm)=(1); gray(1cm)=(0);},
			colorbar style={
			        ytick={0,0.05,...,0.5},
			        yticklabel style={
			            text width=2.5em,
			            align=right,
			            /pgf/number format/.cd,
			                fixed,
			                fixed zerofill
			        }
			    }
			]

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.3850,0.1710) (7.2250,0.3850,0.1710) (7.2250,0.4405,0.1710) (0.4600,0.4405,0.1710)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.4405,0.1080) (7.2250,0.4405,0.1080) (7.2250,0.4860,0.1080) (0.4600,0.4860,0.1080)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.4860,0.0250) (7.2250,0.4860,0.0250) (7.2250,0.4910,0.0250) (0.4600,0.4910,0.0250)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.4910,0.0000) (7.2250,0.4910,0.0000) (7.2250,0.4960,0.0000) (0.4600,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.4960,0.0795) (7.2250,0.4960,0.0795) (7.2250,0.5190,0.0795) (0.4600,0.5190,0.0795)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5190,0.0000) (7.2250,0.5190,0.0000) (7.2250,0.5220,0.0000) (0.4600,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5220,0.0000) (7.2250,0.5220,0.0000) (7.2250,0.5280,0.0000) (0.4600,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5280,0.0035) (7.2250,0.5280,0.0035) (7.2250,0.5410,0.0035) (0.4600,0.5410,0.0035)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5410,0.0000) (7.2250,0.5410,0.0000) (7.2250,0.5455,0.0000) (0.4600,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5455,0.0045) (7.2250,0.5455,0.0045) (7.2250,0.5910,0.0045) (0.4600,0.5910,0.0045)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5910,0.0000) (7.2250,0.5910,0.0000) (7.2250,0.6390,0.0000) (0.4600,0.6390,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.6390,0.0185) (7.2250,0.6390,0.0185) (7.2250,0.6510,0.0185) (0.4600,0.6510,0.0185)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.6510,0.0000) (7.2250,0.6510,0.0000) (7.2250,0.8205,0.0000) (0.4600,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.8205,0.0000) (7.2250,0.8205,0.0000) (7.2250,0.8710,0.0000) (0.4600,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.3850,0.0000) (8.0050,0.3850,0.0000) (8.0050,0.4405,0.0000) (7.2250,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.4405,0.0000) (8.0050,0.4405,0.0000) (8.0050,0.4860,0.0000) (7.2250,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.4860,0.0000) (8.0050,0.4860,0.0000) (8.0050,0.4910,0.0000) (7.2250,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.4910,0.0145) (8.0050,0.4910,0.0145) (8.0050,0.4960,0.0145) (7.2250,0.4960,0.0145)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.4960,0.0000) (8.0050,0.4960,0.0000) (8.0050,0.5190,0.0000) (7.2250,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5190,0.0000) (8.0050,0.5190,0.0000) (8.0050,0.5220,0.0000) (7.2250,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5220,0.0120) (8.0050,0.5220,0.0120) (8.0050,0.5280,0.0120) (7.2250,0.5280,0.0120)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5280,0.0000) (8.0050,0.5280,0.0000) (8.0050,0.5410,0.0000) (7.2250,0.5410,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5410,0.0000) (8.0050,0.5410,0.0000) (8.0050,0.5455,0.0000) (7.2250,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5455,0.0000) (8.0050,0.5455,0.0000) (8.0050,0.5910,0.0000) (7.2250,0.5910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5910,0.0000) (8.0050,0.5910,0.0000) (8.0050,0.6390,0.0000) (7.2250,0.6390,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.6390,0.0000) (8.0050,0.6390,0.0000) (8.0050,0.6510,0.0000) (7.2250,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.6510,0.0000) (8.0050,0.6510,0.0000) (8.0050,0.8205,0.0000) (7.2250,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.8205,0.0000) (8.0050,0.8205,0.0000) (8.0050,0.8710,0.0000) (7.2250,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.3850,0.0000) (8.3500,0.3850,0.0000) (8.3500,0.4405,0.0000) (8.0050,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.4405,0.0000) (8.3500,0.4405,0.0000) (8.3500,0.4860,0.0000) (8.0050,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.4860,0.0000) (8.3500,0.4860,0.0000) (8.3500,0.4910,0.0000) (8.0050,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.4910,0.0000) (8.3500,0.4910,0.0000) (8.3500,0.4960,0.0000) (8.0050,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.4960,0.0000) (8.3500,0.4960,0.0000) (8.3500,0.5190,0.0000) (8.0050,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5190,0.0000) (8.3500,0.5190,0.0000) (8.3500,0.5220,0.0000) (8.0050,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5220,0.0000) (8.3500,0.5220,0.0000) (8.3500,0.5280,0.0000) (8.0050,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5280,0.0455) (8.3500,0.5280,0.0455) (8.3500,0.5410,0.0455) (8.0050,0.5410,0.0455)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5410,0.0000) (8.3500,0.5410,0.0000) (8.3500,0.5455,0.0000) (8.0050,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5455,0.0000) (8.3500,0.5455,0.0000) (8.3500,0.5910,0.0000) (8.0050,0.5910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5910,0.0000) (8.3500,0.5910,0.0000) (8.3500,0.6390,0.0000) (8.0050,0.6390,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.6390,0.0000) (8.3500,0.6390,0.0000) (8.3500,0.6510,0.0000) (8.0050,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.6510,0.0000) (8.3500,0.6510,0.0000) (8.3500,0.8205,0.0000) (8.0050,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.8205,0.0000) (8.3500,0.8205,0.0000) (8.3500,0.8710,0.0000) (8.0050,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.3850,0.0000) (9.1250,0.3850,0.0000) (9.1250,0.4405,0.0000) (8.3500,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.4405,0.0000) (9.1250,0.4405,0.0000) (9.1250,0.4860,0.0000) (8.3500,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.4860,0.0000) (9.1250,0.4860,0.0000) (9.1250,0.4910,0.0000) (8.3500,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.4910,0.0000) (9.1250,0.4910,0.0000) (9.1250,0.4960,0.0000) (8.3500,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.4960,0.0000) (9.1250,0.4960,0.0000) (9.1250,0.5190,0.0000) (8.3500,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5190,0.0150) (9.1250,0.5190,0.0150) (9.1250,0.5220,0.0150) (8.3500,0.5220,0.0150)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5220,0.0000) (9.1250,0.5220,0.0000) (9.1250,0.5280,0.0000) (8.3500,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5280,0.0000) (9.1250,0.5280,0.0000) (9.1250,0.5410,0.0000) (8.3500,0.5410,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5410,0.0000) (9.1250,0.5410,0.0000) (9.1250,0.5455,0.0000) (8.3500,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5455,0.0000) (9.1250,0.5455,0.0000) (9.1250,0.5910,0.0000) (8.3500,0.5910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5910,0.0000) (9.1250,0.5910,0.0000) (9.1250,0.6390,0.0000) (8.3500,0.6390,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.6390,0.0000) (9.1250,0.6390,0.0000) (9.1250,0.6510,0.0000) (8.3500,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.6510,0.0000) (9.1250,0.6510,0.0000) (9.1250,0.8205,0.0000) (8.3500,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.8205,0.0000) (9.1250,0.8205,0.0000) (9.1250,0.8710,0.0000) (8.3500,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.3850,0.0330) (16.5700,0.3850,0.0330) (16.5700,0.4405,0.0330) (9.1250,0.4405,0.0330)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.4405,0.0065) (16.5700,0.4405,0.0065) (16.5700,0.4860,0.0065) (9.1250,0.4860,0.0065)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.4860,0.0210) (16.5700,0.4860,0.0210) (16.5700,0.4910,0.0210) (9.1250,0.4910,0.0210)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.4910,0.0000) (16.5700,0.4910,0.0000) (16.5700,0.4960,0.0000) (9.1250,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.4960,0.0000) (16.5700,0.4960,0.0000) (16.5700,0.5190,0.0000) (9.1250,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5190,0.0000) (16.5700,0.5190,0.0000) (16.5700,0.5220,0.0000) (9.1250,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5220,0.0000) (16.5700,0.5220,0.0000) (16.5700,0.5280,0.0000) (9.1250,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5280,0.0000) (16.5700,0.5280,0.0000) (16.5700,0.5410,0.0000) (9.1250,0.5410,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5410,0.0220) (16.5700,0.5410,0.0220) (16.5700,0.5455,0.0220) (9.1250,0.5455,0.0220)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5455,0.0490) (16.5700,0.5455,0.0490) (16.5700,0.5910,0.0490) (9.1250,0.5910,0.0490)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5910,0.0000) (16.5700,0.5910,0.0000) (16.5700,0.6390,0.0000) (9.1250,0.6390,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.6390,0.0000) (16.5700,0.6390,0.0000) (16.5700,0.6510,0.0000) (9.1250,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.6510,0.0000) (16.5700,0.6510,0.0000) (16.5700,0.8205,0.0000) (9.1250,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.8205,0.0000) (16.5700,0.8205,0.0000) (16.5700,0.8710,0.0000) (9.1250,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.3850,0.0000) (18.8400,0.3850,0.0000) (18.8400,0.4405,0.0000) (16.5700,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.4405,0.0000) (18.8400,0.4405,0.0000) (18.8400,0.4860,0.0000) (16.5700,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.4860,0.0000) (18.8400,0.4860,0.0000) (18.8400,0.4910,0.0000) (16.5700,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.4910,0.0000) (18.8400,0.4910,0.0000) (18.8400,0.4960,0.0000) (16.5700,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.4960,0.0000) (18.8400,0.4960,0.0000) (18.8400,0.5190,0.0000) (16.5700,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5190,0.0000) (18.8400,0.5190,0.0000) (18.8400,0.5220,0.0000) (16.5700,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5220,0.0000) (18.8400,0.5220,0.0000) (18.8400,0.5280,0.0000) (16.5700,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5280,0.0095) (18.8400,0.5280,0.0095) (18.8400,0.5410,0.0095) (16.5700,0.5410,0.0095)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5410,0.0000) (18.8400,0.5410,0.0000) (18.8400,0.5455,0.0000) (16.5700,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5455,0.0295) (18.8400,0.5455,0.0295) (18.8400,0.5910,0.0295) (16.5700,0.5910,0.0295)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5910,0.0260) (18.8400,0.5910,0.0260) (18.8400,0.6390,0.0260) (16.5700,0.6390,0.0260)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.6390,0.0000) (18.8400,0.6390,0.0000) (18.8400,0.6510,0.0000) (16.5700,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.6510,0.1955) (18.8400,0.6510,0.1955) (18.8400,0.8205,0.1955) (16.5700,0.8205,0.1955)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.8205,0.0000) (18.8400,0.8205,0.0000) (18.8400,0.8710,0.0000) (16.5700,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.3850,0.0000) (20.7350,0.3850,0.0000) (20.7350,0.4405,0.0000) (18.8400,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.4405,0.0000) (20.7350,0.4405,0.0000) (20.7350,0.4860,0.0000) (18.8400,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.4860,0.0000) (20.7350,0.4860,0.0000) (20.7350,0.4910,0.0000) (18.8400,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.4910,0.0000) (20.7350,0.4910,0.0000) (20.7350,0.4960,0.0000) (18.8400,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.4960,0.0000) (20.7350,0.4960,0.0000) (20.7350,0.5190,0.0000) (18.8400,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5190,0.0000) (20.7350,0.5190,0.0000) (20.7350,0.5220,0.0000) (18.8400,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5220,0.0000) (20.7350,0.5220,0.0000) (20.7350,0.5280,0.0000) (18.8400,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5280,0.0000) (20.7350,0.5280,0.0000) (20.7350,0.5410,0.0000) (18.8400,0.5410,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5410,0.0000) (20.7350,0.5410,0.0000) (20.7350,0.5455,0.0000) (18.8400,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5455,0.0000) (20.7350,0.5455,0.0000) (20.7350,0.5910,0.0000) (18.8400,0.5910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5910,0.0335) (20.7350,0.5910,0.0335) (20.7350,0.6390,0.0335) (18.8400,0.6390,0.0335)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.6390,0.0000) (20.7350,0.6390,0.0000) (20.7350,0.6510,0.0000) (18.8400,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.6510,0.0000) (20.7350,0.6510,0.0000) (20.7350,0.8205,0.0000) (18.8400,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.8205,0.0300) (20.7350,0.8205,0.0300) (20.7350,0.8710,0.0300) (18.8400,0.8710,0.0300)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.3850,0.0000) (27.7400,0.3850,0.0000) (27.7400,0.4405,0.0000) (20.7350,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.4405,0.0000) (27.7400,0.4405,0.0000) (27.7400,0.4860,0.0000) (20.7350,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.4860,0.0000) (27.7400,0.4860,0.0000) (27.7400,0.4910,0.0000) (20.7350,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.4910,0.0000) (27.7400,0.4910,0.0000) (27.7400,0.4960,0.0000) (20.7350,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.4960,0.0000) (27.7400,0.4960,0.0000) (27.7400,0.5190,0.0000) (20.7350,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5190,0.0000) (27.7400,0.5190,0.0000) (27.7400,0.5220,0.0000) (20.7350,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5220,0.0000) (27.7400,0.5220,0.0000) (27.7400,0.5280,0.0000) (20.7350,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5280,0.0000) (27.7400,0.5280,0.0000) (27.7400,0.5410,0.0000) (20.7350,0.5410,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5410,0.0000) (27.7400,0.5410,0.0000) (27.7400,0.5455,0.0000) (20.7350,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5455,0.0160) (27.7400,0.5455,0.0160) (27.7400,0.5910,0.0160) (20.7350,0.5910,0.0160)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5910,0.0315) (27.7400,0.5910,0.0315) (27.7400,0.6390,0.0315) (20.7350,0.6390,0.0315)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.6390,0.0000) (27.7400,0.6390,0.0000) (27.7400,0.6510,0.0000) (20.7350,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.6510,0.0000) (27.7400,0.6510,0.0000) (27.7400,0.8205,0.0000) (20.7350,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.8205,0.0000) (27.7400,0.8205,0.0000) (27.7400,0.8710,0.0000) (20.7350,0.8710,0.0000)};

			\addplot3[mark=+, draw=none, mark size=0.7, jitter = 0.3] coordinates {
(2.310,0.538,1) (7.070,0.469,1) (7.070,0.469,1) (2.180,0.458,1) (2.180,0.458,1) (2.180,0.458,1) (7.870,0.524,1) (7.870,0.524,1) (7.870,0.524,1) (7.870,0.524,1) (7.870,0.524,1) (7.870,0.524,1) (7.870,0.524,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (5.960,0.499,1) (5.960,0.499,1) (5.960,0.499,1) (5.960,0.499,1) (2.950,0.428,1) (2.950,0.428,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (5.640,0.439,1) (5.640,0.439,1) (5.640,0.439,1) (5.640,0.439,1) (4.000,0.410,1) (1.220,0.403,1) (0.740,0.410,1) (1.320,0.411,1) (5.130,0.453,1) (5.130,0.453,1) (5.130,0.453,1) (5.130,0.453,1) (5.130,0.453,1) (5.130,0.453,1) (1.380,0.416,1) (3.370,0.398,1) (3.370,0.398,1) (6.070,0.409,1) (6.070,0.409,1) (6.070,0.409,1) (10.810,0.413,1) (10.810,0.413,1) (10.810,0.413,1) (10.810,0.413,1) (12.830,0.437,1) (12.830,0.437,1) (12.830,0.437,1) (12.830,0.437,1) (12.830,0.437,1) (12.830,0.437,1) (4.860,0.426,1) (4.860,0.426,1) (4.860,0.426,1) (4.860,0.426,1) (4.490,0.449,1) (4.490,0.449,1) (4.490,0.449,1) (4.490,0.449,1) (3.410,0.489,1) (3.410,0.489,1) (3.410,0.489,1) (3.410,0.489,1) (15.040,0.464,1) (15.040,0.464,1) (15.040,0.464,1) (2.890,0.445,1) (2.890,0.445,1) (2.890,0.445,1) (2.890,0.445,1) (2.890,0.445,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (25.650,0.581,1) (25.650,0.581,1) (25.650,0.581,1) (25.650,0.581,1) (25.650,0.581,1) (25.650,0.581,1) (25.650,0.581,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.871,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (4.050,0.510,1) (4.050,0.510,1) (4.050,0.510,1) (4.050,0.510,1) (4.050,0.510,1) (4.050,0.510,1) (4.050,0.510,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (3.440,0.437,1) (3.440,0.437,1) (3.440,0.437,1) (3.440,0.437,1) (3.440,0.437,1) (3.440,0.437,1) (2.930,0.401,1) (2.930,0.401,1) (0.460,0.422,1) (1.520,0.404,1) (1.520,0.404,1) (1.520,0.404,1) (1.470,0.403,1) (1.470,0.403,1) (2.030,0.415,1) (2.030,0.415,1) (2.680,0.416,1) (2.680,0.416,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (13.890,0.550,1) (13.890,0.550,1) (13.890,0.550,1) (13.890,0.550,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (4.930,0.428,1) (4.930,0.428,1) (4.930,0.428,1) (4.930,0.428,1) (4.930,0.428,1) (4.930,0.428,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (3.640,0.392,1) (3.640,0.392,1) (3.750,0.394,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.575,1) (3.970,0.575,1) (6.960,0.464,1) (6.960,0.464,1) (6.960,0.464,1) (6.960,0.464,1) (6.960,0.464,1) (6.410,0.447,1) (6.410,0.447,1) (6.410,0.447,1) (6.410,0.447,1) (6.410,0.447,1) (3.330,0.443,1) (3.330,0.443,1) (3.330,0.443,1) (3.330,0.443,1) (1.210,0.401,1) (2.970,0.400,1) (2.250,0.389,1) (1.760,0.385,1) (5.320,0.405,1) (5.320,0.405,1) (5.320,0.405,1) (4.950,0.411,1) (4.950,0.411,1) (4.950,0.411,1) (13.920,0.437,1) (13.920,0.437,1) (13.920,0.437,1) (13.920,0.437,1) (13.920,0.437,1) (2.240,0.400,1) (2.240,0.400,1) (2.240,0.400,1) (6.090,0.433,1) (6.090,0.433,1) (6.090,0.433,1) (2.180,0.472,1) (2.180,0.472,1) (2.180,0.472,1) (2.180,0.472,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (3.240,0.460,1) (3.240,0.460,1) (3.240,0.460,1) (6.060,0.438,1) (6.060,0.438,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (1.520,0.442,1) (1.890,0.518,1) (3.780,0.484,1) (3.780,0.484,1) (4.390,0.442,1) (4.390,0.442,1) (4.150,0.429,1) (2.010,0.435,1) (1.250,0.429,1) (1.250,0.429,1) (1.690,0.411,1) (1.690,0.411,1) (2.020,0.410,1) (1.910,0.413,1) (1.910,0.413,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.718,1) (18.100,0.718,1) (18.100,0.718,1) (18.100,0.631,1) (18.100,0.631,1) (18.100,0.631,1) (18.100,0.631,1) (18.100,0.631,1) (18.100,0.668,1) (18.100,0.668,1) (18.100,0.668,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.659,1) (18.100,0.659,1) (18.100,0.597,1) (18.100,0.597,1) (18.100,0.597,1) (18.100,0.597,1) (18.100,0.597,1) (18.100,0.597,1) (18.100,0.693,1) (18.100,0.679,1) (18.100,0.679,1) (18.100,0.679,1) (18.100,0.679,1) (18.100,0.718,1) (18.100,0.718,1) (18.100,0.718,1) (18.100,0.614,1) (18.100,0.614,1) (18.100,0.584,1) (18.100,0.679,1) (18.100,0.584,1) (18.100,0.679,1) (18.100,0.679,1) (18.100,0.679,1) (18.100,0.584,1) (18.100,0.584,1) (18.100,0.584,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.655,1) (18.100,0.655,1) (18.100,0.655,1) (18.100,0.584,1) (18.100,0.580,1) (18.100,0.580,1) (18.100,0.580,1) (18.100,0.532,1) (18.100,0.580,1) (18.100,0.614,1) (18.100,0.584,1) (18.100,0.584,1) (18.100,0.614,1) (18.100,0.614,1) (18.100,0.614,1) (18.100,0.614,1) (18.100,0.532,1) (18.100,0.532,1) (18.100,0.532,1) (18.100,0.532,1) (18.100,0.583,1) (18.100,0.583,1) (18.100,0.583,1) (18.100,0.583,1) (27.740,0.609,1) (27.740,0.609,1) (27.740,0.609,1) (27.740,0.609,1) (27.740,0.609,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (11.930,0.573,1) (11.930,0.573,1) (11.930,0.573,1) (11.930,0.573,1) (11.930,0.573,1) };
		\end{axis}
	\end{tikzpicture}