### 2015-11-22 probability_plot_old/graph_housing_exp1_distr.tex

\begin{tikzpicture}
  \begin{groupplot}[group style={group size=1 by 2}]%,xmin=0,ymin=0,height=4cm,width=5cm,no markers]
  \nextgroupplot[
			view={0}{90},
			enlargelimits=0,
			zmin=0.0, zmax=1.0,
			ylabel=E,
			colorbar right,
    every colorbar/.append style={height=
        2*\pgfkeysvalueof{/pgfplots/parent axis height}+
        \pgfkeysvalueof{/pgfplots/group/vertical sep}},
			point meta min=0, point meta max=0.5,
			colormap={}{ gray(0cm)=(1); gray(1cm)=(0);},
			colorbar style={
			        ytick={0,0.05,...,0.5},
			        yticklabel style={
			            text width=2.5em,
			            align=right,
			            /pgf/number format/.cd,
			                fixed,
			                fixed zerofill
			        }
			    }
			]

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.3850,0.1710) (7.2250,0.3850,0.1710) (7.2250,0.4405,0.1710) (0.4600,0.4405,0.1710)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.4405,0.1080) (7.2250,0.4405,0.1080) (7.2250,0.4860,0.1080) (0.4600,0.4860,0.1080)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.4860,0.0250) (7.2250,0.4860,0.0250) (7.2250,0.4910,0.0250) (0.4600,0.4910,0.0250)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.4910,0.0000) (7.2250,0.4910,0.0000) (7.2250,0.4960,0.0000) (0.4600,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.4960,0.0795) (7.2250,0.4960,0.0795) (7.2250,0.5190,0.0795) (0.4600,0.5190,0.0795)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5190,0.0000) (7.2250,0.5190,0.0000) (7.2250,0.5220,0.0000) (0.4600,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5220,0.0000) (7.2250,0.5220,0.0000) (7.2250,0.5280,0.0000) (0.4600,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5280,0.0035) (7.2250,0.5280,0.0035) (7.2250,0.5410,0.0035) (0.4600,0.5410,0.0035)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5410,0.0000) (7.2250,0.5410,0.0000) (7.2250,0.5455,0.0000) (0.4600,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5455,0.0045) (7.2250,0.5455,0.0045) (7.2250,0.5910,0.0045) (0.4600,0.5910,0.0045)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.5910,0.0000) (7.2250,0.5910,0.0000) (7.2250,0.6390,0.0000) (0.4600,0.6390,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.6390,0.0185) (7.2250,0.6390,0.0185) (7.2250,0.6510,0.0185) (0.4600,0.6510,0.0185)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.6510,0.0000) (7.2250,0.6510,0.0000) (7.2250,0.8205,0.0000) (0.4600,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(0.4600,0.8205,0.0000) (7.2250,0.8205,0.0000) (7.2250,0.8710,0.0000) (0.4600,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.3850,0.0000) (8.0050,0.3850,0.0000) (8.0050,0.4405,0.0000) (7.2250,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.4405,0.0000) (8.0050,0.4405,0.0000) (8.0050,0.4860,0.0000) (7.2250,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.4860,0.0000) (8.0050,0.4860,0.0000) (8.0050,0.4910,0.0000) (7.2250,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.4910,0.0145) (8.0050,0.4910,0.0145) (8.0050,0.4960,0.0145) (7.2250,0.4960,0.0145)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.4960,0.0000) (8.0050,0.4960,0.0000) (8.0050,0.5190,0.0000) (7.2250,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5190,0.0000) (8.0050,0.5190,0.0000) (8.0050,0.5220,0.0000) (7.2250,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5220,0.0120) (8.0050,0.5220,0.0120) (8.0050,0.5280,0.0120) (7.2250,0.5280,0.0120)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5280,0.0000) (8.0050,0.5280,0.0000) (8.0050,0.5410,0.0000) (7.2250,0.5410,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5410,0.0000) (8.0050,0.5410,0.0000) (8.0050,0.5455,0.0000) (7.2250,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5455,0.0000) (8.0050,0.5455,0.0000) (8.0050,0.5910,0.0000) (7.2250,0.5910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.5910,0.0000) (8.0050,0.5910,0.0000) (8.0050,0.6390,0.0000) (7.2250,0.6390,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.6390,0.0000) (8.0050,0.6390,0.0000) (8.0050,0.6510,0.0000) (7.2250,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.6510,0.0000) (8.0050,0.6510,0.0000) (8.0050,0.8205,0.0000) (7.2250,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.2250,0.8205,0.0000) (8.0050,0.8205,0.0000) (8.0050,0.8710,0.0000) (7.2250,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.3850,0.0000) (8.3500,0.3850,0.0000) (8.3500,0.4405,0.0000) (8.0050,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.4405,0.0000) (8.3500,0.4405,0.0000) (8.3500,0.4860,0.0000) (8.0050,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.4860,0.0000) (8.3500,0.4860,0.0000) (8.3500,0.4910,0.0000) (8.0050,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.4910,0.0000) (8.3500,0.4910,0.0000) (8.3500,0.4960,0.0000) (8.0050,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.4960,0.0000) (8.3500,0.4960,0.0000) (8.3500,0.5190,0.0000) (8.0050,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5190,0.0000) (8.3500,0.5190,0.0000) (8.3500,0.5220,0.0000) (8.0050,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5220,0.0000) (8.3500,0.5220,0.0000) (8.3500,0.5280,0.0000) (8.0050,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5280,0.0455) (8.3500,0.5280,0.0455) (8.3500,0.5410,0.0455) (8.0050,0.5410,0.0455)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5410,0.0000) (8.3500,0.5410,0.0000) (8.3500,0.5455,0.0000) (8.0050,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5455,0.0000) (8.3500,0.5455,0.0000) (8.3500,0.5910,0.0000) (8.0050,0.5910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.5910,0.0000) (8.3500,0.5910,0.0000) (8.3500,0.6390,0.0000) (8.0050,0.6390,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.6390,0.0000) (8.3500,0.6390,0.0000) (8.3500,0.6510,0.0000) (8.0050,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.6510,0.0000) (8.3500,0.6510,0.0000) (8.3500,0.8205,0.0000) (8.0050,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.0050,0.8205,0.0000) (8.3500,0.8205,0.0000) (8.3500,0.8710,0.0000) (8.0050,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.3850,0.0000) (9.1250,0.3850,0.0000) (9.1250,0.4405,0.0000) (8.3500,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.4405,0.0000) (9.1250,0.4405,0.0000) (9.1250,0.4860,0.0000) (8.3500,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.4860,0.0000) (9.1250,0.4860,0.0000) (9.1250,0.4910,0.0000) (8.3500,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.4910,0.0000) (9.1250,0.4910,0.0000) (9.1250,0.4960,0.0000) (8.3500,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.4960,0.0000) (9.1250,0.4960,0.0000) (9.1250,0.5190,0.0000) (8.3500,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5190,0.0150) (9.1250,0.5190,0.0150) (9.1250,0.5220,0.0150) (8.3500,0.5220,0.0150)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5220,0.0000) (9.1250,0.5220,0.0000) (9.1250,0.5280,0.0000) (8.3500,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5280,0.0000) (9.1250,0.5280,0.0000) (9.1250,0.5410,0.0000) (8.3500,0.5410,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5410,0.0000) (9.1250,0.5410,0.0000) (9.1250,0.5455,0.0000) (8.3500,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5455,0.0000) (9.1250,0.5455,0.0000) (9.1250,0.5910,0.0000) (8.3500,0.5910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.5910,0.0000) (9.1250,0.5910,0.0000) (9.1250,0.6390,0.0000) (8.3500,0.6390,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.6390,0.0000) (9.1250,0.6390,0.0000) (9.1250,0.6510,0.0000) (8.3500,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.6510,0.0000) (9.1250,0.6510,0.0000) (9.1250,0.8205,0.0000) (8.3500,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(8.3500,0.8205,0.0000) (9.1250,0.8205,0.0000) (9.1250,0.8710,0.0000) (8.3500,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.3850,0.0330) (16.5700,0.3850,0.0330) (16.5700,0.4405,0.0330) (9.1250,0.4405,0.0330)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.4405,0.0065) (16.5700,0.4405,0.0065) (16.5700,0.4860,0.0065) (9.1250,0.4860,0.0065)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.4860,0.0210) (16.5700,0.4860,0.0210) (16.5700,0.4910,0.0210) (9.1250,0.4910,0.0210)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.4910,0.0000) (16.5700,0.4910,0.0000) (16.5700,0.4960,0.0000) (9.1250,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.4960,0.0000) (16.5700,0.4960,0.0000) (16.5700,0.5190,0.0000) (9.1250,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5190,0.0000) (16.5700,0.5190,0.0000) (16.5700,0.5220,0.0000) (9.1250,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5220,0.0000) (16.5700,0.5220,0.0000) (16.5700,0.5280,0.0000) (9.1250,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5280,0.0000) (16.5700,0.5280,0.0000) (16.5700,0.5410,0.0000) (9.1250,0.5410,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5410,0.0220) (16.5700,0.5410,0.0220) (16.5700,0.5455,0.0220) (9.1250,0.5455,0.0220)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5455,0.0490) (16.5700,0.5455,0.0490) (16.5700,0.5910,0.0490) (9.1250,0.5910,0.0490)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.5910,0.0000) (16.5700,0.5910,0.0000) (16.5700,0.6390,0.0000) (9.1250,0.6390,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.6390,0.0000) (16.5700,0.6390,0.0000) (16.5700,0.6510,0.0000) (9.1250,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.6510,0.0000) (16.5700,0.6510,0.0000) (16.5700,0.8205,0.0000) (9.1250,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(9.1250,0.8205,0.0000) (16.5700,0.8205,0.0000) (16.5700,0.8710,0.0000) (9.1250,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.3850,0.0000) (18.8400,0.3850,0.0000) (18.8400,0.4405,0.0000) (16.5700,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.4405,0.0000) (18.8400,0.4405,0.0000) (18.8400,0.4860,0.0000) (16.5700,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.4860,0.0000) (18.8400,0.4860,0.0000) (18.8400,0.4910,0.0000) (16.5700,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.4910,0.0000) (18.8400,0.4910,0.0000) (18.8400,0.4960,0.0000) (16.5700,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.4960,0.0000) (18.8400,0.4960,0.0000) (18.8400,0.5190,0.0000) (16.5700,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5190,0.0000) (18.8400,0.5190,0.0000) (18.8400,0.5220,0.0000) (16.5700,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5220,0.0000) (18.8400,0.5220,0.0000) (18.8400,0.5280,0.0000) (16.5700,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5280,0.0095) (18.8400,0.5280,0.0095) (18.8400,0.5410,0.0095) (16.5700,0.5410,0.0095)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5410,0.0000) (18.8400,0.5410,0.0000) (18.8400,0.5455,0.0000) (16.5700,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5455,0.0295) (18.8400,0.5455,0.0295) (18.8400,0.5910,0.0295) (16.5700,0.5910,0.0295)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.5910,0.0260) (18.8400,0.5910,0.0260) (18.8400,0.6390,0.0260) (16.5700,0.6390,0.0260)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.6390,0.0000) (18.8400,0.6390,0.0000) (18.8400,0.6510,0.0000) (16.5700,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.6510,0.1955) (18.8400,0.6510,0.1955) (18.8400,0.8205,0.1955) (16.5700,0.8205,0.1955)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(16.5700,0.8205,0.0000) (18.8400,0.8205,0.0000) (18.8400,0.8710,0.0000) (16.5700,0.8710,0.0000)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.3850,0.0000) (20.7350,0.3850,0.0000) (20.7350,0.4405,0.0000) (18.8400,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.4405,0.0000) (20.7350,0.4405,0.0000) (20.7350,0.4860,0.0000) (18.8400,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.4860,0.0000) (20.7350,0.4860,0.0000) (20.7350,0.4910,0.0000) (18.8400,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.4910,0.0000) (20.7350,0.4910,0.0000) (20.7350,0.4960,0.0000) (18.8400,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.4960,0.0000) (20.7350,0.4960,0.0000) (20.7350,0.5190,0.0000) (18.8400,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5190,0.0000) (20.7350,0.5190,0.0000) (20.7350,0.5220,0.0000) (18.8400,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5220,0.0000) (20.7350,0.5220,0.0000) (20.7350,0.5280,0.0000) (18.8400,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5280,0.0000) (20.7350,0.5280,0.0000) (20.7350,0.5410,0.0000) (18.8400,0.5410,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5410,0.0000) (20.7350,0.5410,0.0000) (20.7350,0.5455,0.0000) (18.8400,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5455,0.0000) (20.7350,0.5455,0.0000) (20.7350,0.5910,0.0000) (18.8400,0.5910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.5910,0.0335) (20.7350,0.5910,0.0335) (20.7350,0.6390,0.0335) (18.8400,0.6390,0.0335)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.6390,0.0000) (20.7350,0.6390,0.0000) (20.7350,0.6510,0.0000) (18.8400,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.6510,0.0000) (20.7350,0.6510,0.0000) (20.7350,0.8205,0.0000) (18.8400,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(18.8400,0.8205,0.0300) (20.7350,0.8205,0.0300) (20.7350,0.8710,0.0300) (18.8400,0.8710,0.0300)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.3850,0.0000) (27.7400,0.3850,0.0000) (27.7400,0.4405,0.0000) (20.7350,0.4405,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.4405,0.0000) (27.7400,0.4405,0.0000) (27.7400,0.4860,0.0000) (20.7350,0.4860,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.4860,0.0000) (27.7400,0.4860,0.0000) (27.7400,0.4910,0.0000) (20.7350,0.4910,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.4910,0.0000) (27.7400,0.4910,0.0000) (27.7400,0.4960,0.0000) (20.7350,0.4960,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.4960,0.0000) (27.7400,0.4960,0.0000) (27.7400,0.5190,0.0000) (20.7350,0.5190,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5190,0.0000) (27.7400,0.5190,0.0000) (27.7400,0.5220,0.0000) (20.7350,0.5220,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5220,0.0000) (27.7400,0.5220,0.0000) (27.7400,0.5280,0.0000) (20.7350,0.5280,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5280,0.0000) (27.7400,0.5280,0.0000) (27.7400,0.5410,0.0000) (20.7350,0.5410,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5410,0.0000) (27.7400,0.5410,0.0000) (27.7400,0.5455,0.0000) (20.7350,0.5455,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5455,0.0160) (27.7400,0.5455,0.0160) (27.7400,0.5910,0.0160) (20.7350,0.5910,0.0160)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.5910,0.0315) (27.7400,0.5910,0.0315) (27.7400,0.6390,0.0315) (20.7350,0.6390,0.0315)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.6390,0.0000) (27.7400,0.6390,0.0000) (27.7400,0.6510,0.0000) (20.7350,0.6510,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.6510,0.0000) (27.7400,0.6510,0.0000) (27.7400,0.8205,0.0000) (20.7350,0.8205,0.0000)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(20.7350,0.8205,0.0000) (27.7400,0.8205,0.0000) (27.7400,0.8710,0.0000) (20.7350,0.8710,0.0000)};

			\addplot3[mark=+, draw=none, mark size=0.7] coordinates {
(2.310,0.538,1) (7.070,0.469,1) (7.070,0.469,1) (2.180,0.458,1) (2.180,0.458,1) (2.180,0.458,1) (7.870,0.524,1) (7.870,0.524,1) (7.870,0.524,1) (7.870,0.524,1) (7.870,0.524,1) (7.870,0.524,1) (7.870,0.524,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (8.140,0.538,1) (5.960,0.499,1) (5.960,0.499,1) (5.960,0.499,1) (5.960,0.499,1) (2.950,0.428,1) (2.950,0.428,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (6.910,0.448,1) (5.640,0.439,1) (5.640,0.439,1) (5.640,0.439,1) (5.640,0.439,1) (4.000,0.410,1) (1.220,0.403,1) (0.740,0.410,1) (1.320,0.411,1) (5.130,0.453,1) (5.130,0.453,1) (5.130,0.453,1) (5.130,0.453,1) (5.130,0.453,1) (5.130,0.453,1) (1.380,0.416,1) (3.370,0.398,1) (3.370,0.398,1) (6.070,0.409,1) (6.070,0.409,1) (6.070,0.409,1) (10.810,0.413,1) (10.810,0.413,1) (10.810,0.413,1) (10.810,0.413,1) (12.830,0.437,1) (12.830,0.437,1) (12.830,0.437,1) (12.830,0.437,1) (12.830,0.437,1) (12.830,0.437,1) (4.860,0.426,1) (4.860,0.426,1) (4.860,0.426,1) (4.860,0.426,1) (4.490,0.449,1) (4.490,0.449,1) (4.490,0.449,1) (4.490,0.449,1) (3.410,0.489,1) (3.410,0.489,1) (3.410,0.489,1) (3.410,0.489,1) (15.040,0.464,1) (15.040,0.464,1) (15.040,0.464,1) (2.890,0.445,1) (2.890,0.445,1) (2.890,0.445,1) (2.890,0.445,1) (2.890,0.445,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (8.560,0.520,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (10.010,0.547,1) (25.650,0.581,1) (25.650,0.581,1) (25.650,0.581,1) (25.650,0.581,1) (25.650,0.581,1) (25.650,0.581,1) (25.650,0.581,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (21.890,0.624,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.871,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.871,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (19.580,0.605,1) (4.050,0.510,1) (4.050,0.510,1) (4.050,0.510,1) (4.050,0.510,1) (4.050,0.510,1) (4.050,0.510,1) (4.050,0.510,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (2.460,0.488,1) (3.440,0.437,1) (3.440,0.437,1) (3.440,0.437,1) (3.440,0.437,1) (3.440,0.437,1) (3.440,0.437,1) (2.930,0.401,1) (2.930,0.401,1) (0.460,0.422,1) (1.520,0.404,1) (1.520,0.404,1) (1.520,0.404,1) (1.470,0.403,1) (1.470,0.403,1) (2.030,0.415,1) (2.030,0.415,1) (2.680,0.416,1) (2.680,0.416,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (10.590,0.489,1) (13.890,0.550,1) (13.890,0.550,1) (13.890,0.550,1) (13.890,0.550,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.504,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (6.200,0.507,1) (4.930,0.428,1) (4.930,0.428,1) (4.930,0.428,1) (4.930,0.428,1) (4.930,0.428,1) (4.930,0.428,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (5.860,0.431,1) (3.640,0.392,1) (3.640,0.392,1) (3.750,0.394,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.647,1) (3.970,0.575,1) (3.970,0.575,1) (6.960,0.464,1) (6.960,0.464,1) (6.960,0.464,1) (6.960,0.464,1) (6.960,0.464,1) (6.410,0.447,1) (6.410,0.447,1) (6.410,0.447,1) (6.410,0.447,1) (6.410,0.447,1) (3.330,0.443,1) (3.330,0.443,1) (3.330,0.443,1) (3.330,0.443,1) (1.210,0.401,1) (2.970,0.400,1) (2.250,0.389,1) (1.760,0.385,1) (5.320,0.405,1) (5.320,0.405,1) (5.320,0.405,1) (4.950,0.411,1) (4.950,0.411,1) (4.950,0.411,1) (13.920,0.437,1) (13.920,0.437,1) (13.920,0.437,1) (13.920,0.437,1) (13.920,0.437,1) (2.240,0.400,1) (2.240,0.400,1) (2.240,0.400,1) (6.090,0.433,1) (6.090,0.433,1) (6.090,0.433,1) (2.180,0.472,1) (2.180,0.472,1) (2.180,0.472,1) (2.180,0.472,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (9.900,0.544,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (7.380,0.493,1) (3.240,0.460,1) (3.240,0.460,1) (3.240,0.460,1) (6.060,0.438,1) (6.060,0.438,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (5.190,0.515,1) (1.520,0.442,1) (1.890,0.518,1) (3.780,0.484,1) (3.780,0.484,1) (4.390,0.442,1) (4.390,0.442,1) (4.150,0.429,1) (2.010,0.435,1) (1.250,0.429,1) (1.250,0.429,1) (1.690,0.411,1) (1.690,0.411,1) (2.020,0.410,1) (1.910,0.413,1) (1.910,0.413,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.770,1) (18.100,0.718,1) (18.100,0.718,1) (18.100,0.718,1) (18.100,0.631,1) (18.100,0.631,1) (18.100,0.631,1) (18.100,0.631,1) (18.100,0.631,1) (18.100,0.668,1) (18.100,0.668,1) (18.100,0.668,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.671,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.700,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.693,1) (18.100,0.659,1) (18.100,0.659,1) (18.100,0.597,1) (18.100,0.597,1) (18.100,0.597,1) (18.100,0.597,1) (18.100,0.597,1) (18.100,0.597,1) (18.100,0.693,1) (18.100,0.679,1) (18.100,0.679,1) (18.100,0.679,1) (18.100,0.679,1) (18.100,0.718,1) (18.100,0.718,1) (18.100,0.718,1) (18.100,0.614,1) (18.100,0.614,1) (18.100,0.584,1) (18.100,0.679,1) (18.100,0.584,1) (18.100,0.679,1) (18.100,0.679,1) (18.100,0.679,1) (18.100,0.584,1) (18.100,0.584,1) (18.100,0.584,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.740,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.713,1) (18.100,0.655,1) (18.100,0.655,1) (18.100,0.655,1) (18.100,0.584,1) (18.100,0.580,1) (18.100,0.580,1) (18.100,0.580,1) (18.100,0.532,1) (18.100,0.580,1) (18.100,0.614,1) (18.100,0.584,1) (18.100,0.584,1) (18.100,0.614,1) (18.100,0.614,1) (18.100,0.614,1) (18.100,0.614,1) (18.100,0.532,1) (18.100,0.532,1) (18.100,0.532,1) (18.100,0.532,1) (18.100,0.583,1) (18.100,0.583,1) (18.100,0.583,1) (18.100,0.583,1) (27.740,0.609,1) (27.740,0.609,1) (27.740,0.609,1) (27.740,0.609,1) (27.740,0.609,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (9.690,0.585,1) (11.930,0.573,1) (11.930,0.573,1) (11.930,0.573,1) (11.930,0.573,1) (11.930,0.573,1) };

    \nextgroupplot[
			view={0}{90},
			enlargelimits=0,
			zmin=0.0, zmax=1.0,
			xlabel=E,
			ylabel=K,
			%colorbar,
			point meta min=0, point meta max=0.5,
			colormap={}{ gray(0cm)=(1); gray(1cm)=(0);}
			%colorbar style={
			 %       ytick={0,0.05,...,0.5},
			  %      yticklabel style={
			   %         text width=2.5em,
			    %        align=right,
			     %       /pgf/number format/.cd,
			      %          fixed,
			       %         fixed zerofill
			        %}
			    %}
			]

    		\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(3.5610,2.9000,0.0035) (3.7120,2.9000,0.0035) (3.7120,51.4000,0.0035) (3.5610,51.4000,0.0035)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(3.5610,51.4000,0.0015) (3.7120,51.4000,0.0015) (3.7120,77.7500,0.0015) (3.5610,77.7500,0.0015)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(3.5610,77.7500,0.0050) (3.7120,77.7500,0.0050) (3.7120,100.0000,0.0050) (3.5610,100.0000,0.0050)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(3.7120,2.9000,0.1580) (6.5425,2.9000,0.1580) (6.5425,51.4000,0.1580) (3.7120,51.4000,0.1580)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(3.7120,51.4000,0.1440) (6.5425,51.4000,0.1440) (6.5425,77.7500,0.1440) (3.7120,77.7500,0.1440)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(3.7120,77.7500,0.4130) (6.5425,77.7500,0.4130) (6.5425,100.0000,0.4130) (3.7120,100.0000,0.4130)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(6.5425,2.9000,0.0950) (7.4370,2.9000,0.0950) (7.4370,51.4000,0.0950) (6.5425,51.4000,0.0950)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(6.5425,51.4000,0.0550) (7.4370,51.4000,0.0550) (7.4370,77.7500,0.0550) (6.5425,77.7500,0.0550)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(6.5425,77.7500,0.0710) (7.4370,77.7500,0.0710) (7.4370,100.0000,0.0710) (6.5425,100.0000,0.0710)};

			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.4370,2.9000,0.0350) (8.7800,2.9000,0.0350) (8.7800,51.4000,0.0350) (7.4370,51.4000,0.0350)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.4370,51.4000,0.0085) (8.7800,51.4000,0.0085) (8.7800,77.7500,0.0085) (7.4370,77.7500,0.0085)};
			\addplot3[patch,shader=faceted,patch type=rectangle] coordinates {(7.4370,77.7500,0.0105) (8.7800,77.7500,0.0105) (8.7800,100.0000,0.0105) (7.4370,100.0000,0.0105)};

			\addplot3[mark=+, draw=none, mark size=0.7] coordinates {
(6.575,65.200,1) (6.421,78.900,1) (7.185,61.100,1) (6.998,45.800,1) (7.147,54.200,1) (6.430,58.700,1) (6.012,66.600,1) (6.172,96.100,1) (5.631,100.000,1) (6.004,85.900,1) (6.377,94.300,1) (6.009,82.900,1) (5.889,39.000,1) (5.949,61.800,1) (6.096,84.500,1) (5.834,56.500,1) (5.935,29.300,1) (5.990,81.700,1) (5.456,36.600,1) (5.727,69.500,1) (5.570,98.100,1) (5.965,89.200,1) (6.142,91.700,1) (5.813,100.000,1) (5.924,94.100,1) (5.599,85.700,1) (5.813,90.300,1) (6.047,88.800,1) (6.495,94.400,1) (6.674,87.300,1) (5.713,94.100,1) (6.072,100.000,1) (5.950,82.000,1) (5.701,95.000,1) (6.096,96.900,1) (5.933,68.200,1) (5.841,61.400,1) (5.850,41.500,1) (5.966,30.200,1) (6.595,21.800,1) (7.024,15.800,1) (6.770,2.900,1) (6.169,6.600,1) (6.211,6.500,1) (6.069,40.000,1) (5.682,33.800,1) (5.786,33.300,1) (6.030,85.500,1) (5.399,95.300,1) (5.602,62.000,1) (5.963,45.700,1) (6.115,63.000,1) (6.511,21.100,1) (5.998,21.400,1) (5.888,47.600,1) (7.249,21.900,1) (6.383,35.700,1) (6.816,40.500,1) (6.145,29.200,1) (5.927,47.200,1) (5.741,66.200,1) (5.966,93.400,1) (6.456,67.800,1) (6.762,43.400,1) (7.104,59.500,1) (6.290,17.800,1) (5.787,31.100,1) (5.878,21.400,1) (5.594,36.800,1) (5.885,33.000,1) (6.417,6.600,1) (5.961,17.500,1) (6.065,7.800,1) (6.245,6.200,1) (6.273,6.000,1) (6.286,45.000,1) (6.279,74.500,1) (6.140,45.800,1) (6.232,53.700,1) (5.874,36.600,1) (6.727,33.500,1) (6.619,70.400,1) (6.302,32.200,1) (6.167,46.700,1) (6.389,48.000,1) (6.630,56.100,1) (6.015,45.100,1) (6.121,56.800,1) (7.007,86.300,1) (7.079,63.100,1) (6.417,66.100,1) (6.405,73.900,1) (6.442,53.600,1) (6.211,28.900,1) (6.249,77.300,1) (6.625,57.800,1) (6.163,69.600,1) (8.069,76.000,1) (7.820,36.900,1) (7.416,62.500,1) (6.727,79.900,1) (6.781,71.300,1) (6.405,85.400,1) (6.137,87.400,1) (6.167,90.000,1) (5.851,96.700,1) (5.836,91.900,1) (6.127,85.200,1) (6.474,97.100,1) (6.229,91.200,1) (6.195,54.400,1) (6.715,81.600,1) (5.913,92.900,1) (6.092,95.400,1) (6.254,84.200,1) (5.928,88.200,1) (6.176,72.500,1) (6.021,82.600,1) (5.872,73.100,1) (5.731,65.200,1) (5.870,69.700,1) (6.004,84.100,1) (5.961,92.900,1) (5.856,97.000,1) (5.879,95.800,1) (5.986,88.400,1) (5.613,95.600,1) (5.693,96.000,1) (6.431,98.800,1) (5.637,94.700,1) (6.458,98.900,1) (6.326,97.700,1) (6.372,97.900,1) (5.822,95.400,1) (5.757,98.400,1) (6.335,98.200,1) (5.942,93.500,1) (6.454,98.400,1) (5.857,98.200,1) (6.151,97.900,1) (6.174,93.600,1) (5.019,100.000,1) (5.403,100.000,1) (5.468,100.000,1) (4.903,97.800,1) (6.130,100.000,1) (5.628,100.000,1) (4.926,95.700,1) (5.186,93.800,1) (5.597,94.900,1) (6.122,97.300,1) (5.404,100.000,1) (5.012,88.000,1) (5.709,98.500,1) (6.129,96.000,1) (6.152,82.600,1) (5.272,94.000,1) (6.943,97.400,1) (6.066,100.000,1) (6.510,100.000,1) (6.250,92.600,1) (7.489,90.800,1) (7.802,98.200,1) (8.375,93.900,1) (5.854,91.800,1) (6.101,93.000,1) (7.929,96.200,1) (5.877,79.200,1) (6.319,96.100,1) (6.402,95.200,1) (5.875,94.600,1) (5.880,97.300,1) (5.572,88.500,1) (6.416,84.100,1) (5.859,68.700,1) (6.546,33.100,1) (6.020,47.200,1) (6.315,73.400,1) (6.860,74.400,1) (6.980,58.400,1) (7.765,83.300,1) (6.144,62.200,1) (7.155,92.200,1) (6.563,95.600,1) (5.604,89.800,1) (6.153,68.800,1) (7.831,53.600,1) (6.782,41.100,1) (6.556,29.100,1) (7.185,38.900,1) (6.951,21.500,1) (6.739,30.800,1) (7.178,26.300,1) (6.800,9.900,1) (6.604,18.800,1) (7.875,32.000,1) (7.287,34.100,1) (7.107,36.600,1) (7.274,38.300,1) (6.975,15.300,1) (7.135,13.900,1) (6.162,38.400,1) (7.610,15.700,1) (7.853,33.200,1) (8.034,31.900,1) (5.891,22.300,1) (6.326,52.500,1) (5.783,72.700,1) (6.064,59.100,1) (5.344,100.000,1) (5.960,92.100,1) (5.404,88.600,1) (5.807,53.800,1) (6.375,32.300,1) (5.412,9.800,1) (6.182,42.400,1) (5.888,56.000,1) (6.642,85.100,1) (5.951,93.800,1) (6.373,92.400,1) (6.951,88.500,1) (6.164,91.300,1) (6.879,77.700,1) (6.618,80.800,1) (8.266,78.300,1) (8.725,83.000,1) (8.040,86.500,1) (7.163,79.900,1) (7.686,17.000,1) (6.552,21.400,1) (5.981,68.100,1) (7.412,76.900,1) (8.337,73.300,1) (8.247,70.400,1) (6.726,66.500,1) (6.086,61.500,1) (6.631,76.500,1) (7.358,71.600,1) (6.481,18.500,1) (6.606,42.200,1) (6.897,54.300,1) (6.095,65.100,1) (6.358,52.900,1) (6.393,7.800,1) (5.593,76.500,1) (5.605,70.200,1) (6.108,34.900,1) (6.226,79.200,1) (6.433,49.100,1) (6.718,17.500,1) (6.487,13.000,1) (6.438,8.900,1) (6.957,6.800,1) (8.259,8.400,1) (6.108,32.000,1) (5.876,19.100,1) (7.454,34.200,1) (8.704,86.900,1) (7.333,100.000,1) (6.842,100.000,1) (7.203,81.800,1) (7.520,89.400,1) (8.398,91.500,1) (7.327,94.500,1) (7.206,91.600,1) (5.560,62.800,1) (7.014,84.600,1) (8.297,67.000,1) (7.470,52.600,1) (5.920,61.500,1) (5.856,42.100,1) (6.240,16.300,1) (6.538,58.700,1) (7.691,51.800,1) (6.758,32.900,1) (6.854,42.800,1) (7.267,49.000,1) (6.826,27.600,1) (6.482,32.100,1) (6.812,32.200,1) (7.820,64.500,1) (6.968,37.200,1) (7.645,49.700,1) (7.923,24.800,1) (7.088,20.800,1) (6.453,31.900,1) (6.230,31.500,1) (6.209,31.300,1) (6.315,45.600,1) (6.565,22.900,1) (6.861,27.900,1) (7.148,27.700,1) (6.630,23.400,1) (6.127,18.400,1) (6.009,42.300,1) (6.678,31.100,1) (6.549,51.000,1) (5.790,58.000,1) (6.345,20.100,1) (7.041,10.000,1) (6.871,47.400,1) (6.590,40.400,1) (6.495,18.400,1) (6.982,17.700,1) (7.236,41.100,1) (6.616,58.100,1) (7.420,71.900,1) (6.849,70.300,1) (6.635,82.500,1) (5.972,76.700,1) (4.973,37.800,1) (6.122,52.800,1) (6.023,90.400,1) (6.266,82.800,1) (6.567,87.300,1) (5.705,77.700,1) (5.914,83.200,1) (5.782,71.700,1) (6.382,67.200,1) (6.113,58.800,1) (6.426,52.300,1) (6.376,54.300,1) (6.041,49.900,1) (5.708,74.300,1) (6.415,40.100,1) (6.431,14.700,1) (6.312,28.900,1) (6.083,43.700,1) (5.868,25.800,1) (6.333,17.200,1) (6.144,32.200,1) (5.706,28.400,1) (6.031,23.300,1) (6.316,38.100,1) (6.310,38.500,1) (6.037,34.500,1) (5.869,46.300,1) (5.895,59.600,1) (6.059,37.300,1) (5.985,45.400,1) (5.968,58.500,1) (7.241,49.300,1) (6.540,59.700,1) (6.696,56.400,1) (6.874,28.100,1) (6.014,48.500,1) (5.898,52.300,1) (6.516,27.700,1) (6.635,29.700,1) (6.939,34.500,1) (6.490,44.400,1) (6.579,35.900,1) (5.884,18.500,1) (6.728,36.100,1) (5.663,21.900,1) (5.936,19.500,1) (6.212,97.400,1) (6.395,91.000,1) (6.127,83.400,1) (6.112,81.300,1) (6.398,88.000,1) (6.251,91.100,1) (5.362,96.200,1) (5.803,89.000,1) (8.780,82.900,1) (3.561,87.900,1) (4.963,91.400,1) (3.863,100.000,1) (4.970,100.000,1) (6.683,96.800,1) (7.016,97.500,1) (6.216,100.000,1) (5.875,89.600,1) (4.906,100.000,1) (4.138,100.000,1) (7.313,97.900,1) (6.649,93.300,1) (6.794,98.800,1) (6.380,96.200,1) (6.223,100.000,1) (6.968,91.900,1) (6.545,99.100,1) (5.536,100.000,1) (5.520,100.000,1) (4.368,91.200,1) (5.277,98.100,1) (4.652,100.000,1) (5.000,89.500,1) (4.880,100.000,1) (5.390,98.900,1) (5.713,97.000,1) (6.051,82.500,1) (5.036,97.000,1) (6.193,92.600,1) (5.887,94.700,1) (6.471,98.800,1) (6.405,96.000,1) (5.747,98.900,1) (5.453,100.000,1) (5.852,77.800,1) (5.987,100.000,1) (6.343,100.000,1) (6.404,100.000,1) (5.349,96.000,1) (5.531,85.400,1) (5.683,100.000,1) (4.138,100.000,1) (5.608,100.000,1) (5.617,97.900,1) (6.852,100.000,1) (5.757,100.000,1) (6.657,100.000,1) (4.628,100.000,1) (5.155,100.000,1) (4.519,100.000,1) (6.434,100.000,1) (6.782,90.800,1) (5.304,89.100,1) (5.957,100.000,1) (6.824,76.500,1) (6.411,100.000,1) (6.006,95.300,1) (5.648,87.600,1) (6.103,85.100,1) (5.565,70.600,1) (5.896,95.400,1) (5.837,59.700,1) (6.202,78.700,1) (6.193,78.100,1) (6.380,95.600,1) (6.348,86.100,1) (6.833,94.300,1) (6.425,74.800,1) (6.436,87.900,1) (6.208,95.000,1) (6.629,94.600,1) (6.461,93.300,1) (6.152,100.000,1) (5.935,87.900,1) (5.627,93.900,1) (5.818,92.400,1) (6.406,97.200,1) (6.219,100.000,1) (6.485,100.000,1) (5.854,96.600,1) (6.459,94.800,1) (6.341,96.400,1) (6.251,96.600,1) (6.185,98.700,1) (6.417,98.300,1) (6.749,92.600,1) (6.655,98.200,1) (6.297,91.800,1) (7.393,99.300,1) (6.728,94.100,1) (6.525,86.500,1) (5.976,87.900,1) (5.936,80.300,1) (6.301,83.700,1) (6.081,84.400,1) (6.701,90.000,1) (6.376,88.400,1) (6.317,83.000,1) (6.513,89.900,1) (6.209,65.400,1) (5.759,48.200,1) (5.952,84.700,1) (6.003,94.500,1) (5.926,71.000,1) (5.713,56.700,1) (6.167,84.000,1) (6.229,90.700,1) (6.437,75.000,1) (6.980,67.600,1) (5.427,95.400,1) (6.162,97.400,1) (6.484,93.600,1) (5.304,97.300,1) (6.185,96.700,1) (6.229,88.000,1) (6.242,64.700,1) (6.750,74.900,1) (7.061,77.000,1) (5.762,40.300,1) (5.871,41.900,1) (6.312,51.900,1) (6.114,79.800,1) (5.905,53.200,1) (5.454,92.700,1) (5.414,98.300,1) (5.093,98.000,1) (5.983,98.800,1) (5.983,83.500,1) (5.707,54.000,1) (5.926,42.600,1) (5.670,28.800,1) (5.390,72.900,1) (5.794,70.600,1) (6.019,65.300,1) (5.569,73.500,1) (6.027,79.700,1) (6.593,69.100,1) (6.120,76.700,1) (6.976,91.000,1) (6.794,89.300,1) (6.030,80.800,1) };
  \end{groupplot}
\end{tikzpicture}